---
title: "Labelled Data for Form Languages"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Labelled Data for Form Languages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  echo = TRUE,
  comment = "#>"
)

```

Projects in `Kobotoolbox` supports multiple languages. This can be utilized to label questions and choices values in either single or multiple-choice questions.

`robotoolbox` enables you to access labels in various languages in your form. The [`labelled`](https://cran.r-project.org/package=labelled/vignettes/intro_labelled.html) package is used to encode variable labels for question labels and value labels for choice labels.

To demonstrate its application, data will be loaded from the following project that employs three languages: English, French, and Arabic.

### Survey questions

| type             | name      | label::English (en)  | label::Francais (fr)               | label::Arabic (ar) |
|:-----------------|:----------|:---------------------|:-----------------------------------|:-------------------|
| start            | start     |                      |                                    |                    |
| end              | end       |                      |                                    |                    |
| today            | today     |                      |                                    |                    |
| text             | full_name | What is your name?   | Quel est votre nom ?               | ما اسمك ؟          |
| select_one yesno | pet_yesno | Do you have any pet? | Avez-vous un animal de compagnie ? | هل تمتلك حيوانا أليفا ؟ |


### Choices

| list_name | name | label::English (en) | label::Francais (fr) | label::Arabic (ar) |
|:----------|-----:|:--------------------|:---------------------|:-------------------|
| yesno     |    1 | Yes                 | Oui                  | نعم                |
| yesno     |    0 | No                  | Non                  | لا                 |
|           |      |                     |                      |                    |


### Loading the project asset

This project, which was uploaded to a `Kobotoolbox` server, can be loaded into your session using its unique identifier `uid`.

```{r, eval = FALSE}
library(robotoolbox)
library(dplyr)
uid <- "aYuTZn9vegi3Z49MXwKjep"
asset <- kobo_asset(uid)
asset
```

```{r, echo = FALSE}
library(robotoolbox)
library(dplyr)
asset <- asset_ml
asset
```

### Listing available languages

The `kobo_lang` function can be used to list all languages available in the project.

```{r list_lang}
kobo_lang(asset)
```

### Reading data in each languages

The `kobo_data` function allows the use of the `lang` parameter to specify the language to be loaded. The spelling of each language can be checked first using `kobo_lang`.

```{r, eval = FALSE}
df_en <- kobo_data(asset, lang = "English (en)")
df_fr <- kobo_data(asset, lang = "Francais (fr)")
df_ar <- kobo_data(asset, lang = "Arabic (ar)")
glimpse(df_en)
```

```{r, echo = FALSE}
df_en <- data_ml_en
df_fr <- data_ml_fr
df_ar <- data_ml_ar
glimpse(df_en)
```

When the parameter isn't specified, the default language is used, which in this project is `English (en)`.

```{r, eval = FALSE}
df_default <- kobo_data(asset)
all.equal(df_default, df_en)
```

```{r, echo = FALSE}
df_default <- data_ml_default
all.equal(df_default, df_en)
```

### Accessing variable labels

The `var_label` function from the `labelled` package can be used to access question labels.

```{r var_label1}
library(labelled)
var_label(df_en$full_name)
var_label(df_fr$full_name)
var_label(df_ar$full_name)
```


```{r var_label2}
var_label(df_en$pet_yesno)
var_label(df_fr$pet_yesno)
var_label(df_ar$pet_yesno)
```

### Accessing `select_one` type question labels

The `to_factor` function can convert the values of single-choice questions into labels.

```{r val_label}
table(to_factor(df_en$pet_yesno))
table(to_factor(df_fr$pet_yesno))
table(to_factor(df_ar$pet_yesno))
```

When `character` is preferred over `factor`, `to_character` serves as an alternative to `to_factor`.

```{r val_label_character}
count(df_ar, pet_yesno_ar = to_character(pet_yesno))
```

All of these functions come from the `labelled` package; you can explore this package further through its [documentation](https://larmarange.github.io/labelled/).
