---
title: "Loading Data with Repeating Groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Loading Data with Repeating Groups}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  echo = TRUE,
  comment = "#>"
)
```

```{r setup}
library(robotoolbox)
library(dplyr)
library(dm)
```

## Loading data

```{r asset_list_fake, eval = FALSE}
l <- kobo_asset_list()
l
## # A tibble: 19 × 7
##    uid     name     asset_type owner_username date_created
##    <chr>   <chr>    <chr>      <chr>          <dttm>
##  1 aANhxw… nested_… survey     xxxxx         2022-01-05 21:22:50
##  2 aUmFLX… Sample … survey     xxxxx         2022-01-06 12:41:31
##  3 ascwAw… nested_… survey     xxxxx         2022-01-05 21:22:22
##  4 an8Awk… Clone o… survey     xxxxx         2022-01-04 11:52:22
##  5 a4QDRG… nested_… survey     xxxxx         2022-01-03 15:41:52
##  6 aNYcey… Clone o… survey     xxxxx         2021-12-31 14:33:16
##  7 aZSRWs… test_re… survey     xxxxx         2021-12-30 11:44:04
##  8 akGLub… complic… survey     xxxxx         2021-12-03 14:34:32
##  9 aALrrW… Pakista… survey     xxxxx         2021-11-29 07:55:28
## 10 a5NMnC… duplica… survey     xxxxx         2021-11-18 10:44:17
## 11 aNyj7z… Enhanci… survey     xxxx          2021-09-22 11:50:06
## 12 ajwxzH… GBV sur… survey     xxxxx         2021-07-27 11:36:16
## 13 aa9iC3… Sample … survey     xxxxx         2021-10-12 14:10:18
## 14 aMVqE7… abupbcy… survey     xxxxx         2021-07-27 11:36:06
## 15 az75no… Evaluat… survey     xxxxx         2021-06-21 14:09:27
## 16 aEnSVD… test_re… survey     xxxxx         2021-06-22 14:00:20
## 17 a7rSJQ… P21 key… survey     xxxxx         2021-04-30 11:51:41
## 18 aSxBDW… p21 hou… survey     xxxxx         2021-05-07 10:24:42
## 19 ano26o… Test pe… survey     xxxxx         2021-02-04 08:52:51
## # … with 2 more variables: date_modified <dttm>,
## #   submissions <int>
```


```{r asset_list, echo = FALSE}
l <- kobo_asset_list()
```

### Survey questions

| type                 | name              | label::English (en)                              | label::Francais (fr)                                  | repeat_count    | calculation                                    |
|:---------------------|:------------------|:-------------------------------------------------|:------------------------------------------------------|:----------------|:-----------------------------------------------|
| start                | start             |                                                  |                                                       |                 |                                                |
| end                  | end               |                                                  |                                                       |                 |                                                |
| today                | today             |                                                  |                                                       |                 |                                                |
| begin_repeat         | demo              | Demographic Characteristics                      | Caracteristique Demographique                         |                 |                                                |
| text                 | name              | Name                                             | Nom                                                   |                 |                                                |
| integer              | age               | Age                                              | Age                                                   |                 |                                                |
| select_one sex       | sex               | Sex                                              | Sexe                                                  |                 |                                                |
| integer              | hobby             | How many hobbies does \${name} have?              | Combien de hobbies \${name} a ?                        |                 |                                                |
| select_one yesno     | morelang          | Does \${name} speak more than one language?       | Est-ce que \${name} parle plus d'une langue ?          |                 |                                                |
| calculate            | name_individual   |                                                  |                                                       |                 | indexed-repeat(\${name}, \${demo}, position(..)) |
| begin_repeat         | hobbies_list      | List of Hobbies                                  | Liste de hobbies                                      | \${hobby}        |                                                |
| text                 | hobbies           | Hobbies of \${name_individual}                    | Hobbies de \${name_individual}                         |                 |                                                |
| end_repeat           |                   |                                                  |                                                       |                 |                                                |
| begin_repeat         | lang_list         | List of Languages                                | Liste de langues                                      | \${morelang}     |                                                |
| select_multiple lang | langs             | Languages spoken by \${name_individual}           | Langue parle par \${name_individual}                   |                 |                                                |
| end_repeat           |                   |                                                  |                                                       |                 |                                                |
| end_repeat           |                   |                                                  |                                                       |                 |                                                |
| calculate            | family_count      |                                                  |                                                       |                 | count(\${demo})                                 |
| note                 | family_count_note | Number of family members: \${family_count}        | Nombre de membre dans la famille:  \${family_count}    |                 |                                                |
| begin_repeat         | education         | Education information                            | Information sur l'education                           | \${family_count} |                                                |
| calculate            | name_individual2  |                                                  |                                                       |                 | indexed-repeat(\${name}, \${demo}, position(..)) |
| select_one edu_level | edu_level         | What is \${name_individual2}'s level of education | Quel est le niveau d'education de \${name_individual2} |                 |                                                |
| end_repeat           |                   |                                                  |                                                       |                 |                                                |



### Choices

|list_name | name|label::English (en)      |label::Francais (fr) |
|:---------|----:|:------------------------|:--------------------|
|sex       |    1|Male                     |Homme                |
|sex       |    2|Female                   |Femme                |
|sex       |    3|Prefer not to say        |Prefere ne pas dire  |
|edu_level |    1|Primary                  |Primaire             |
|edu_level |    2|Secondary                |Secondaire           |
|edu_level |    3|Higher Secondary & Above |Lycee et superieur   |
|yesno     |    1|Yes                      |Oui                  |
|yesno     |    0|No                       |Non                  |
|lang      |    1|French                   |Francais             |
|lang      |    2|Spanish                  |Espagnol             |
|lang      |    3|Arabic                   |Arabe                |
|lang      |   99|Other                    |Autre                |


### Loading the project asset

```{r asset}
uid <- filter(l, name == "nested_roster") |>
  pull(uid)
asset <- kobo_asset(uid)
asset
```

### Extracting the data

```{r read_data}
df <- kobo_data(asset)
df
```

## Manipulating repeat groups as dm object

###  Visualizing the relationship between tables

```{r draw_data}
dm_draw(df)
```

### Number of rows of each table

```{r nrow_data}
dm_nrow(df)
```

### A `dm` object is a list of `tibble`

```{r access_specific1}
glimpse(df$main)
```

```{r access_specific2}
glimpse(df$education)
```

## Filtering data


```{r filter_data}
df |>
  dm_filter("main", `_index` == 2) |>
  dm_apply_filters() |>
  dm_nrow()
```

## Joining tables


```{r join_two}
df |>
  dm_join_to_tbl("main",
                 "education",
                 join = left_join) |>
  glimpse()
```


```{r join_all}
df |>
  dm_flatten_to_tbl(start = "hobbies_list",
                    join = left_join) |>
  glimpse()
```
