---
title: "Loading Data with Repeating Groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Loading Data with Repeating Groups}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  echo = TRUE,
  comment = "#>"
)
```

`Kobotoolbox` supports questions that can be grouped to be answered repeatedly. A typical use case is when implementing a household survey, with some series of questions for each household member.

## Loading data

In `Kobotoolbox`, this feature is implemented by repeating a group of questions. It uses the concept of [`repeat group`](https://support.kobotoolbox.org/group_repeat.html). It works by inserting the questions you want to repeat inside a `begin_repeat`/`end_repeat` loop. `repeat group` can be nested, making it possible to have a group of questions repeated within another `repeat group` instance. It can be illustrated with the following project and associated form.


```{r setup, echo = FALSE}
library(robotoolbox)
library(dplyr)
library(dm)
```


```{r asset_list, echo = FALSE}
l <- asset_list
```

### Survey questions

| type                 | name              | label::English (en)                              | label::Francais (fr)                                  | repeat_count    | calculation                                    |
|:---------------------|:------------------|:-------------------------------------------------|:------------------------------------------------------|:----------------|:-----------------------------------------------|
| start                | start             |                                                  |                                                       |                 |                                                |
| end                  | end               |                                                  |                                                       |                 |                                                |
| today                | today             |                                                  |                                                       |                 |                                                |
| **begin_repeat**         | demo              | Demographic Characteristics                      | Caracteristique Demographique                         |                 |                                                |
| text                 | name              | Name                                             | Nom                                                   |                 |                                                |
| integer              | age               | Age                                              | Age                                                   |                 |                                                |
| select_one sex       | sex               | Sex                                              | Sexe                                                  |                 |                                                |
| integer              | hobby             | How many hobbies does \${name} have?              | Combien de hobbies \${name} a ?                        |                 |                                                |
| select_one yesno     | morelang          | Does \${name} speak more than one language?       | Est-ce que \${name} parle plus d'une langue ?          |                 |                                                |
| calculate            | name_individual   |                                                  |                                                       |                 | indexed-repeat(\${name}, \${demo}, position(..)) |
| **begin_repeat**         | hobbies_list      | List of Hobbies                                  | Liste de hobbies                                      | \${hobby}        |                                                |
| text                 | hobbies           | Hobbies of \${name_individual}                    | Hobbies de \${name_individual}                         |                 |                                                |
| **end_repeat**           |                   |                                                  |                                                       |                 |                                                |
| **begin_repeat**         | lang_list         | List of Languages                                | Liste de langues                                      | \${morelang}     |                                                |
| select_multiple lang | langs             | Languages spoken by \${name_individual}           | Langue parle par \${name_individual}                   |                 |                                                |
| **end_repeat**           |                   |                                                  |                                                       |                 |                                                |
| **end_repeat**           |                   |                                                  |                                                       |                 |                                                |
| calculate            | family_count      |                                                  |                                                       |                 | count(\${demo})                                 |
| note                 | family_count_note | Number of family members: \${family_count}        | Nombre de membre dans la famille:  \${family_count}    |                 |                                                |
| **begin_repeat**         | education         | Education information                            | Information sur l'education                           | \${family_count} |                                                |
| calculate            | name_individual2  |                                                  |                                                       |                 | indexed-repeat(\${name}, \${demo}, position(..)) |
| select_one edu_level | edu_level         | What is \${name_individual2}'s level of education | Quel est le niveau d'education de \${name_individual2} |                 |                                                |
| **end_repeat**           |                   |                                                  |                                                       |                 |                                                |



### Choices

|list_name | name|label::English (en)      |label::Francais (fr) |
|:---------|----:|:------------------------|:--------------------|
|sex       |    1|Male                     |Homme                |
|sex       |    2|Female                   |Femme                |
|sex       |    3|Prefer not to say        |Prefere ne pas dire  |
|edu_level |    1|Primary                  |Primaire             |
|edu_level |    2|Secondary                |Secondaire           |
|edu_level |    3|Higher Secondary & Above |Lycee et superieur   |
|yesno     |    1|Yes                      |Oui                  |
|yesno     |    0|No                       |Non                  |
|lang      |    1|French                   |Francais             |
|lang      |    2|Spanish                  |Espagnol             |
|lang      |    3|Arabic                   |Arabe                |
|lang      |   99|Other                    |Autre                |


### Loading the project asset

The above form was uploaded to the server, it's named `nested_roster`, and can be loaded from the list of asset `asset_list`.

```{r, eval = FALSE}
library(robotoolbox)
library(dplyr)
asset_list <- kobo_asset_list()
uid <- filter(asset_list, name == "nested_roster") |>
  pull(uid)
asset <- kobo_asset(uid)
asset
```

```{r, echo = FALSE}
asset <- asset_rg
asset
```

### Extracting the data

We can see that the output is not a standard a `data.frame`. It's a listing of each `repeat group` loop in our form.

```{r, eval = FALSE}
df <- kobo_data(asset)
df
```

```{r, echo = FALSE}
df <- data_rg
df
```

The output is a `dm` object, from the the `dm` package.

```{r, echo = FALSE}
class(df)
```

## Manipulating `repeat group` as `dm` object

A `dm` object is a list of linked `data.frame`, and it can be manipulated using the `dm` package.

###  Visualizing the relationship between tables

In order to understand, how it's stored we can visualize the schema and relationship between tables (repeat groups loop). The schema can be drawn using the the `dm_draw` function.

```{r draw_data}
library(dm)
dm_draw(df)
```

### Number of rows of each table

The `dm` package has a lot of helper functions to manipulated `dm` object. We can get the number of rows of each table using `dm_nrow`

```{r nrow_data}
dm_nrow(df)
```

### A `dm` object is a list of `data.frame`

A `dm` object is a `list`. As in any list of `data.frame`, you can extract each table, and analyze it separately.
The principal table, where you have the first `repeat group` is named `main`.

```{r access_specific1}
glimpse(df$main)
```

The other tables uses their form names, for example `education` for the education `repeat group`.

```{r access_specific2}
glimpse(df$education)
```

## Filtering data

One major advantage of using the `dm` package, is the ability to dynamically filter tables while keeping the links between them.
In this example, filtering the `main` table will propagate to the `education` and `demo` tables. The `hobbies_list` and `lang_list` tables are linked to the `demo` table, and will be filtered too.

```{r filter_data}
df |>
  dm_filter(main = (`_index` == 2)) |>
  dm_nrow()
```

## Joining tables

In some cases, it's simpler to analyze the joined data, `dm_flatten_to_tbl` can be used for safely joining the data keeping its structure and the links between tables.
We can join the `education` table to the `main` table, by usign `dm_flatten_to_tbl` and starting from `education`.

```{r join_two}
df |>
  dm_flatten_to_tbl(.start = education,
                    .join = left_join) |>
  glimpse()
```

We can push this logic to have the widest table possible through a cascade of joins from a deeper table (`start` argument) to the `main` table. Using `hobbies_list` as an example, it'll perform two joins.
`hobbies_list` will be joined to the `demo` table, and the `demo` table will be joined to the `main` table.

```{r join_all}
df |>
  dm_flatten_to_tbl(.start = hobbies_list,
                    .join = left_join,
                    .recursive = TRUE) |>
  glimpse()
```

You can learn a lot about the [`dm`](https://cran.r-project.org/package=dm) package going through its extensive documentation.
