---
title: "Geographic data collection and spatial data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Geographic data collection and spatial data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  echo = TRUE,
  comment = "#>"
)
```

Using `Kobotoolbox`, we can collect spatial data using three question types: **geopoint**, **geotrace** and **geoshape**.

**Geopoint:**<br>
A geopoint question type captures a single geographic coordinate (latitude and longitude) along with altitude and accuracy. This is useful for marking locations, such as houses, schools, or water sources.

**Geotrace:**<br>
A geotrace question type collects a series of connected geographic coordinates, forming a line. This can be used to map routes, paths, or boundaries.

**Geoshape:**<br>
A geoshape question type captures a series of geographic coordinates that form a closed polygon. This is useful for defining areas, such as land parcels, agricultural fields, or protected zones.

In order to use these data we need to parse them in a GIS friendly format. `robotoolbox` uses [Well-Known Text (WKT)](https://libgeos.org/specifications/wkt/), a standard markup language for representing vector geometry, to represent points (`geopoint`), lines (`geotrace`) and polygons (`geoshape`).


## Spatial data

The form below is a toy example to showcase how spatial field types are mapped using `robotoolbox`.

```{r setup, echo = FALSE}
library(robotoolbox)
library(dplyr)
library(sf)
library(mapview)
mapview::mapviewOptions(
    fgb = FALSE,
    basemaps = c(
      "Esri.WorldImagery",
      "Esri.WorldShadedRelief",
      "OpenTopoMap",
      "OpenStreetMap"),
    layers.control.pos = "topright")
```


```{r asset_list, echo = FALSE}
l <- asset_list
```

### Survey questions

|name                |type     |label                          |
|:-------------------|:--------|:------------------------------|
|point               |geopoint |Record a location              |
|point_description   |text     |Describe the recorded location |
|line                |geotrace |Record a line                  |
|line_description    |text     |Describe the recorded line     |
|polygon             |geoshape |Record a polygon               |
|polygon_description |text     |Describe the recorded polygon  |

We have 3 columns of spatial types: `point`, `line` and `polygon`.

### Loading the project

The above form was uploaded to the server, it's the only project named `Spatial data`, and can be loaded from the list of assets `asset_list`.

```{r, eval = FALSE}
library(robotoolbox)
library(dplyr)
asset_list <- kobo_asset_list()
uid <- filter(asset_list, name == "Spatial data") |>
  pull(uid)
asset <- kobo_asset(uid)
asset
```

```{r, echo = FALSE}
asset <- asset_spatial
asset
```

We have just one submission, we recorded one location using `geopoint`, used `geotrace` to map a portion of a road, and stadium was outlined with the `geoshape` question type.

### Extracting the data

From the assets, we can now extract the submissions.

```{r, eval = FALSE}
df <- kobo_data(asset)
glimpse(df)
```

```{r, echo = FALSE}
df <- data_spatial
glimpse(df)
```

We can see that we have all of our three columns `point`, `line` and `polygon`. For each columns, we have a WKT column.

```{r}
pull(df, point)
pull(df, point_wkt)
```

For `geopoint`, `robotoolbox` also provides the latitude, longitude, altitude and precision columns.


```{r}
df |>
  select(starts_with("point_"))
```

WKT column for `geotrace` question type for the `line` column.

```{r}
pull(df, line)
pull(df, line_wkt)
```

WKT column for `geoshape` question type for the `polygon` column.

```{r}
pull(df, polygon)
pull(df, polygon_wkt)
```

Now that we know how `robotoolbox` store spatial question types, we can now turn these column to spatial object `R` can use for spatial data analysis.

## Geopoint

The standard way to manipulate spatial vector data in `R` is to use the `sf` package.
`sf` stands for Simple Features and it extends a data.frame by adding a geometry list-column.
It provides an interface to [GDAL](http://www.gdal.org/), [GEOS](https://libgeos.org/), [PRØJ](http://proj.org/) and [S2](https://s2geometry.io/) to manipulate and visualize spatial data.

It's easy to create an `sf` object from a column containing [`WKT` representation of geometry](https://libgeos.org/specifications/wkt/).

```{r}
point_sf <- st_as_sf(select(data_spatial, point_wkt, point_description),
                     wkt = "point_wkt", crs = 4326)
mapview(point_sf)
```

## Geotrace

We can also transform Kobotoolbox `geotrace` question to an `sf` line object, using the `WKT` column `line_wkt`.

```{r}
line_sf <- st_as_sf(select(data_spatial, line_wkt, line_description),
                     wkt = "line_wkt", crs = 4326)
mapview(line_sf)
```


## Geoshape

The column `polygon_wkt` can be used to create an `sf` polygon object.

```{r}
poly_sf <- st_as_sf(select(data_spatial, polygon_wkt, polygon_description),
                    wkt = "polygon_wkt", crs = 4326)
mapview(poly_sf)
```


You can learn a lot about the `sf` packages and spatial data analysis with R from the excellent [Geocomputation with R](https://r.geocompx.org/) and the `sf` [package documentation](https://r-spatial.github.io/sf/).
